---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wbpip

<!-- badges: start -->

<!-- badges: end -->

The goal of `{wbpip}` is to twofold. First it contains the methodological
formulations currently used by the World Bank to estimate and monitor global
poverty and to calculate cross-country and time comparable socioeconomic
measures. Second, it provides a suit of user-friendly functions that allows
anybody to estimate poverty and inequality measures with their own data, using
the methodological approach of the World Bank.

## Installation

[*Not available as of now (`r Sys.Date()`)*]{style="color:red"}

<!-- You can install the released version of wbpip from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->

<!-- install.packages("wbpip") -->

<!-- ``` -->

<!-- And the development version from [GitHub](https://github.com/) with: -->

<!-- ``` r -->

<!-- # install.packages("devtools") -->

<!-- devtools::install_github("PIP-Technical-Team/wbpip") -->

<!-- ``` -->

# Basic end-user workflows

Basic workflows are based on single data files (micro, grouped, bin, or
synthetic data). The assumption for these workflows is that they can be
satisfied by feeding a data file to different functions.

Functions in `{wbpip}` are divided in two main groups: computational functions
and end-user functions. Computational functions are the internal functions of
the package that do the calculations. End-user functions are wrappers of the
computational functions that fit well in data science workflows such as the
tidyverse.

Most computational functions follow this convention: dd_compute_ss(), where
**dd** stands for data type and **ss** stands from statistic to compute. Data
type could **md** for microdata or **gd** from grouped data. Even though there
are others type of data such as bin or synthetic data, the computational
procedure is the same as of micro data. The statistic to compute is
self-explanatory. It could be poverty, gini, mean log deviation (mld), among
others. Other computational functions have the prefix `prod_`, which refers to
production. These functions make use of lower-level computational functions to
create production tables used in the Poverty and Inequality Platform.

End-user functions are prefixed with `get_`. These functions are intended to be
used as part of any data science workflow using pipes like the tidyverse. This
is why, the first argument of these functions is the data frame with the data.
For convenience, some functions return one single stat, whereas others return a
compendium of related stats:

## Single stats

-   `get_poverty_rate(df, povline)`
    `compute_poverty_rate(welfare, weight, povline)`
-   `get_number_poor(df, povline, pop)`
-   `get_poverty_gap(df, povline)`
-   `get_poverty_severity(df, povline)`
-   `get_watts_index(df, povline)`
-   `get_mean(df)`
-   `get_median(df)`
-   `get_gini(df)`
-   `get_polarization(df)`
-   `get_mld(df)`
-   `get_quantile(n = 10)`

# Multiple stats

-   `get_distributional_stats()`
-   `get_poverty_stats()`

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(wbpip)
## basic example code
```
