---
title: "review_select_lorenz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{review_select_lorenz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wbpip)
```

## Background
Povcalnet uses two methods to estimate poverty and inequality statistics from 
grouped data.  

- One method is based on fitting a Lorenz Quadratic functional form to the 
grouped data  
- the other one uses a Beta Lorenz function  

Povcalnet then choose the statistics to be returned based on a set of rules.

This vignette focuses on the higher level `gd_compute_pip_stats()` function that
handles the application of both functional forms and the selection of the final
results.

## High level example
```{r}
# Input definition
welfare_mean    <- 51.56
ppp             <- 3.69
daily_povline   <- 1.9
monthly_povline <- daily_povline * 365 / 12

# Create grouped data (Type 1)
# http://iresearch.worldbank.org/povcalnet/PovCalculator.aspx
population <- c(0.0005,
                0.0032,
                0.014799999999999999,
                0.0443,
                0.0991,
                0.257,
                0.4385,
                0.5938,
                0.7089,
                1)

welfare <- c(5.824760527229386e-05,
             0.000604029410841011,
             0.0037949334793616948,
             0.013988878652244477,
             0.036992164583098786,
             0.12140708906131342,
             0.24531391873082081,
             0.37446670169288321,
             0.48753116241194566,
             1)


# Estimate poverty statistics
out <- wbpip:::gd_compute_pip_stats(welfare         = welfare,
                                    population      = population,
                                    requested_mean  = welfare_mean,
                                    povline         = monthly_povline,
                                    default_ppp     = ppp)

out

```
